Comparing corrosion control treatments for drinking water using a robust Bayesian generalized additive model

### 2022-07-19 Rmarkdown/paper-si.Rmd
PLC specs, figure annotations, edits to intro

flows %>% 
  rename_all(str_to_sentence) %>% 
  knitr::kable(caption = "**Table S3.** Pipe array flow schedule; sampling occurred at T4, and FY-001, -002, and -003 are defined in Figure S1.")

>>>>

flows %>% 
  rename_all(str_to_sentence) %>% 
  knitr::kable(caption = "**Table S3.** Pipe array flow schedule; sampling occurred at T4, and FY-001, -002, and -003 are defined in Figure S1. A Modicon M221 logic controller (Schneider Electric, Canada) was used for process control. Three 2-inch ASCO 8210G100 solenoid valves (i.e., FY-001, -002 and -003) were wired to the logic controller, opening and closing according to the schedule.")

### 2022-07-19 Rmarkdown/paper.Rmd
PLC specs, figure annotations, edits to intro

(1) Non-linear time trends and seasonal variation. These features can be difficult to describe without over- or under-fitting. They can also make it difficult to identify a stable phase, as typical guidance for analyzing pipe loop study data suggests [@wysock_statistical_1995]. 
(2) Left-censored values (measurements below the detection limit), which many popular statistical tests are not designed to handle.
(3) Autocorrelation, or memory effects. Autocorrelation means the current state of the system depends on its history. For example, the effect of a perturbation to corrosion scale is evident for some time after the fact. From a statistical perspective, autocorrelation makes parameter estimates more uncertain [@helsel1992statistical] by reducing the effective sample size (i.e., the information content of the sample) [@gelman_bayesian_2014]. It is very likely to be a feature of any pipe loop study dataset where effluent from pipes is sampled repeatedly over time.
(4) An irregular sampling frequency, which makes it more difficult to estimate the degree of autocorrelation.
(5) Extreme lead or copper concentrations due to particle release. Extreme values are often inconsistent with the common assumption of  Gaussian residuals.

@@

Left-censoring and autocorrelation can be particularly challenging to model. Left-censoring is often addressed by substituting a constant for values below the detection limit and sometimes by removing the censored values. Neither approach is advisable because both introduce artificial patterns that can distort results, sometimes seriously [@helsel_statistics_2012]. Autocorrelation overstates the precision of statistical estimates [@gelman_bayesian_2014]. As a consequence, it inflates the probability of falsely rejecting null hypotheses [@helsel1992statistical]. The effect of autocorrelation can be surprisingly large: 100 observations from a time series may provide no greater precision than 10 independent observations if the 100 are strongly correlated in time [@helsel1992statistical]. That is, *n* autocorrelated observations provide less information than *n* independent observations, and the difference can be large if autocorrelation is strong [@gelman_bayesian_2014]. Accounting for autocorrelation can improve predictive accuracy, but if it is ignored, analysts run the risk of making incorrect decisions based on the data. And while rank-based procedures such as the rank-sum test can accommodate left-censoring, most are not designed for autocorrelated data [@helsel1992statistical; @helsel_statistics_2012].

@@

Flow through the pipe sections was programmed to approximate typical use and is summarized in Table S3. Samples were collected weekly until Mar. 2021, after which they were collected every two weeks. At each sampling event, 1 L was collected from each pipe after a 6-hour stagnation period. Samples were collected at a flow rate of approximately 3 L min^-1^.

>>>>

(1) Non-linear time trends and seasonal variation. These features can be difficult to describe without over- or under-fitting. They can also make it difficult to identify a stable phase, as typical guidance for analyzing pipe loop study data suggests [@wysock_statistical_1995]. 
(2) Left-censored values (measurements below the detection limit), which many popular statistical tests are not designed to handle.
(3) Autocorrelation, or memory effects. Autocorrelation means the current state of the system depends on its history. For example, the effect of a perturbation to corrosion scale is evident for some time after the fact. From a statistical perspective, autocorrelation makes parameter estimates more uncertain [@helsel_statistical] by reducing the effective sample size (i.e., the information content of the sample) [@gelman_bayesian_2014]. It is very likely to be a feature of any pipe loop study dataset where effluent from pipes is sampled repeatedly over time.
(4) An irregular sampling frequency, which makes it more difficult to estimate the degree of autocorrelation.
(5) Extreme lead or copper concentrations due to particle release. Extreme values are often inconsistent with the common assumption of  Gaussian residuals.

@@

Left-censoring and autocorrelation can be particularly challenging to model. Left-censoring is often addressed by substituting a constant for values below the detection limit and sometimes by removing the censored values. Neither approach is advisable because both introduce artificial patterns that can distort results, sometimes seriously [@helsel_statistics_2012]. Autocorrelation overstates the precision of statistical estimates [@gelman_bayesian_2014]. As a consequence, it inflates the probability of falsely rejecting null hypotheses [@helsel_statistical]. The effect of autocorrelation can be surprisingly large: 100 observations from a time series may provide no greater precision than 10 independent observations if the 100 are strongly correlated in time [@helsel_statistical]. That is, *n* autocorrelated observations provide less information than *n* independent observations, and the difference can be large if autocorrelation is strong [@gelman_bayesian_2014]. Accounting for autocorrelation can improve predictive accuracy, but if it is ignored, analysts run the risk of making incorrect decisions based on the data. And while rank-based procedures such as the rank-sum test can accommodate left-censoring, most are not designed for autocorrelated data [@helsel_statistical; @helsel_statistics_2012]. Likewise, basic parametric methods cannot describe many pipe loop datasets adequately: the paired t-test assumes random sampling of the measurement pairs [@illowsky_introductory_2018; @helsel_statistical] and simple repeated measures ANOVA has a limited capacity to model autocorrelation [@raadt_pressing_2019]. Neither test accommodates censoring and both assume normality.

@@

Flow through the pipe sections was programmed to approximate typical use, and details are available in Table S3. Samples were collected weekly until Mar. 2021, after which they were collected every two weeks. At each sampling event, 1 L was collected from each pipe after a 6-hour stagnation period. Samples were collected at a flow rate of approximately 3 L min^-1^.
